Assessing \Htwo{} and \COtwo{} Network Strategies in Europe

### 2023-12-12 report/submissions/journal/paper/paper.tex
paper: add discussed changes in Intro

The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors. While management of both \COtwo{} and \Htwo{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructure effectively interacts with carbon management technologies, including carbon capture, transport, storage, utilization, and sequestration. Hydrogen offers an efficient way to transport and store energy over long distances and timeframes, and can be produced from electrolysis or steam methane reforming (SMR). Carbon can be stored in geological formations to prevent it from entering the atmosphere, a process known as carbon sequestration (CS). Additionally, carbon is a crucial material input for carbonaceous fuels needed for aviation (synthetic kerosene from Fischer-Tropsch synthesis), shipping (synthetic methanol) and industrial feedstocks (e.g., synthetic methane). From hereon, we jointly refer to the production of these carbonaceous fuels as Carbon Utilization (CU). Carbon can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, or synthetic carbonaceous fuels through point-source carbon capture (CC) techniques, or harvested from the atmosphere using direct air capture (DAC).

@@

To our knowledge, no study has yet considered the co-optimization and comprehensive assessment of both \COtwo{} and \Htwo{} networks in a fully sector-coupled energy system. However, we argue that such an assessment is strongly needed to avoid suboptimal investments and to identify synergies between hydrogen and carbon management technologies. In this paper, we present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies. The study is conducted using the PyPSA-Eur energy system model and encompasses all relevant energy sectors. We explore competition between \COtwo{} and \Htwo{} networks by adding them separately. Our evaluation focuses on the transport of \COtwo{} and \Htwo{} through their respective networks on the European continent. We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon networks varies based on different levels of available sequestration potential.

>>>>

The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors. While management of both \COtwo{} and \Htwo{}, also refered to as carbon and hydrogen, is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructure effectively interacts with carbon management technologies, including carbon capture, transport, storage, utilization, and sequestration. Hydrogen offers an efficient way to transport and store energy over long distances and timeframes, and can be produced from electrolysis or steam methane reforming (SMR), for example. Carbon can be stored in geological formations to prevent it from entering the atmosphere, a process known as carbon sequestration (CS). Additionally, carbon can be combined with hydrogen to produce carbonaceous fuels needed for aviation (synthetic kerosene from Fischer-Tropsch synthesis), shipping (synthetic methanol) and industrial feedstocks (e.g., synthetic methane). From hereon, we jointly refer to the production of these carbonaceous fuels as Carbon Utilization (CU). Carbon can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, or synthetic carbonaceous fuels through point-source carbon capture (CC) techniques, or harvested from the atmosphere using direct air capture (DAC). In combination, \COtwo{} and \Htwo{} networks are crucial to enable these carbon management strategies necessary for net-zero economies.

@@

To our knowledge, no study has yet considered the co-optimization and comprehensive assessment of both \COtwo{} and \Htwo{} networks in a fully sector-coupled energy system. However, we argue that such an assessment is strongly needed to avoid suboptimal investments and to identify synergies between hydrogen and carbon management technologies. In this paper, we present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies. The study is conducted using the PyPSA-Eur energy system model and encompasses all relevant energy sectors. We investigate the competitive transport dynamics between \COtwo{} and \Htwo{} networks. This is done by contrasting models that feature only one of these networks, neither, or both combined. We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon and hydrogen networks varies given net-zero or net-negative emissions targets.
